Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of neutron star - black hole binary observations

### 2016-09-01 Paper/paper.tex
updated fig 7 to corner plots

We construct NSBH waveforms using the LEA+ model. We inject these into zero
noise and run an MCMC sampler on the injections using BBH templates (LEA+ with tidal terms
set to zero). We perform this analysis for a variety of mass ratios
$q\in\{2,3,4,5\}$, i.e. $\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$,
and BH spins $\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. The signal
strength is varied from $\rho=20\rightarrow70$. Our choice of this SNR range is
motivated by previous studies which suggest that signatures of NS tidal effects
become observable only at relatively large SNRs~\cite{Lackey:2013axa}, and is validated
by results that follow. At design sensitivity, if we expect $10-300$ NSBH detections 
a year~\cite{Abadie:2010cfa}, and if BH masses are uniformly likely up to
$\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we would expect to detect $0.3-7$ disruptive
NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
year with $\rho\geq 70$. For our priors, we choose uniform distribution for 
both component masses as well as for black hole spins, i.e. $\mbh\in[1.2,25]M_\odot$,
$\mns\in[1.2,3]M_\odot$ and $-0.75\leq \chibh\leq +0.75$.
% \prayush{FIXME}

@@

\textbf{for signals with different $\lambdans$ values (shown in legend).}\par\medskip

@@

\includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
\includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
\includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
\includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
\includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
\includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
\includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
\includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
\includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}

@@

%
%Ultimately, the extent of tidal moduilations of emitted GWs depends on the
%equation of state of neutron star matter, which becomes an indirect
%measurable in GW astrophysics.
% 
%We therefore turn to the question: If we do observe GWs from disruptive 
%coalescing NSBHs, what information can we extract from them about the internal
%structure of neutron stars?
% 
We begin by calculating the accuracy with which we measure
$\lambdans$ from {\it single} GW observations. We sample the same set of
disruptive NSBH systems as in the previous section, i.e. those with
$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$.
For each unique combination of these parameters, we generate separate 
zero-noise injections at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model. We
perform full Bayesian analyses on each of these injections using LEA+ templates,
and obtain probabilistic estimates for source binary parameters using GW information
alone (including the NS tidal deformability parameter $\lambdans$). Our priors
on component masses and spins remain as in the previous section, while $\lambdans$
is sampled uniformly from $[0,4000]$.

>>>>

We construct NSBH waveforms using the LEA+ model. We inject these into zero
noise and run an MCMC sampler on the injections using BBH templates (LEA+ with tidal terms
set to zero). We perform this analysis for a variety of mass ratios
$q\in\{2,3,4,5\}$, i.e. $\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$,
and BH spins $\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. The signal
strength is varied from $\rho=20\rightarrow70$. Our choice of this SNR range is
motivated by previous studies which suggest that signatures of NS tidal effects
become observable only at relatively large SNRs~\cite{Lackey:2013axa}, and is validated
by results that follow. At design sensitivity, if we expect $10-300$ NSBH detections 
a year~\cite{Abadie:2010cfa}, and if BH masses are uniformly likely up to
$\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we would expect to detect $0.3-7$ disruptive
NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
year with $\rho\geq 70$. For our priors, we choose uniform distributions for 
both component masses, and black hole spin: $\mbh\in[1.2,25]M_\odot$;
$\mns\in[1.2,3]M_\odot$; and $-0.75\leq \chibh\leq +0.75$.
% \prayush{FIXME}

@@

\textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
from left to right.}\par\medskip

@@

\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}

@@

%
%Ultimately, the extent of tidal moduilations of emitted GWs depends on the
%equation of state of neutron star matter, which becomes an indirect
%measurable in GW astrophysics.
% 
%We therefore turn to the question: If we do observe GWs from disruptive 
%coalescing NSBHs, what information can we extract from them about the internal
%structure of neutron stars?
% 
We begin by calculating the accuracy with which we measure
$\lambdans$ from {\it single} GW observations. We sample the same set of
disruptive NSBH systems as in the previous section, i.e. those with
$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$.
For each unique combination of these parameters, we generate separate 
zero-noise injections at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model. We
perform full Bayesian analyses on each of these injections using LEA+ templates,
and obtain probabilistic estimates for source binary parameters using GW information
alone (including the NS tidal deformability parameter $\lambdans$). Our priors
on component masses and spins remain as in the previous section, while $\lambdans$
is sampled uniformly from $[0,4000]$, and mass-ratio is additionally restricted to
$2\leq q\leq 6$.
